Solutions to Integer Programming from Quantum Annealing

### 2020-09-04 paper/main.tex
comments

\begin{abstract}
 While quantum computing proposes promising solutions to computational problems not accessible with classical approaches, due to current hardware constraints, most quantum algorithms are not yet capable of computing systems of practical relevance, and classical counterparts outperform them.
 To practically benefit from quantum architecture, one has to identify problems and algorithms with favorable scaling and improve on corresponding limitations depending on available hardware.
 For this reason, we developed an algorithm that solves integer linear programming problems (ILP), a classically NP-hard problem, on a quantum annealer (QA), and investigated problem and hardware-specific limitations.
 This work presents the formalism of how to map ILP problems to the annealing architectures.
 It illustrates the effects of decorrelation and many-body localization, for the minimum dominating set problem, and compares annealing results against numerical simulations of the quantum architecture.
 We find that the QA ILP algorithm outperforms random guessing but is limited to small problems and that annealing schedules can be adjusted to reduce the effects of decorrelation--which is qualitatively confirmed by simulations.
 {\color{red}Furthermore we find evidence for Many-body localization effects ... and find that FC decoherence is more relevant than local decoherence...}

@@

Integer linear programming (ILP) is an integer optimization problem subject to inequality constraints
\begin{align}
 \label{eq:initial-ip-def}
 \vec x_0 = \mathrm{arg} & \min\limits_{x}(\sum_i c_i x_i)
\end{align}
subject to
\begin{align}
 \label{eq:ilp-constraints}
  & \sum_i A_{ai}x_i +b_a \leq 0 \,, \\
  & x_i  \in \mathbbm{N}\, ,
\end{align}
where the natural numbers $\mathbbm{N}$ include zero, $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations.

@@

An important ILP is the Minimum Dominating Set problem which is represantivly considered in this work.
For a given a graph $G(E,V)$, defined by the set of $V$ vertices and $E$ edges, and a dominating set $D$ is a specific subset of vertices $D \subseteq V$.
In particluar $D$ is a dominating set if all vertices in $V$ but not in $D$ are adjacent to at least one vertex in $D$.
This is equivalent to requiring the set of nearest-neighbor vertices of $D$ (exclusive) and $D$ cover all vertices $N(D) \cup D = V$ (an example is given by Fig.~\ref{fig:dominating_sets}a).
The set $D$ is a minimal dominating set if there is no proper subset of $D$ that is a dominating set, {\it{i.e.}}, the removal of any vertex in $D$ results in $N(D) \cup D  \neq V$.
An example is given by Fig.~\ref{fig:dominating_sets}b.
The domination number of $D$ is given by the cardinality of $|D| \equiv \overline{\overline{D}}$.
The minimum dominating set (MDS) is defined by $D$ with the smallest domination number.
Fig.~\ref{fig:dominating_sets}c shows an example of the minimum dominating set of $G(V, E)$ and is different from the minimal dominating set.
We emphasize that while the maximum independent set is always a minimal dominating set as exemplified by Fig.~\ref{fig:dominating_sets}b, the minimum dominating set, in general, can have a smaller domination number.
As a result, the solution to the dominating set problem can not be obtained by solving for the maximum independent set, a problem that is well studied for quantum annealers~\cite{}.

@@

In this paper, we propose a new algorithm that maps ILP problems to QUBOs.
This is achieved by the introduction of slack variables $s$ which turn the inequalities eq.~\eqref{eq:ilp-constraints} to equalities
\begin{align}
  & \sum_i A_{a i}x_i + s_a + b_a = 0, \\
  & s_a \in \mathbb{N},                \\
  & x_i \in \mathbb{N}.
\end{align}
As a result, we can restrict both $x_i$ and $s_i$ to sample over only positive integer values.

@@

If the annealer does not reach thermal equilibrium, for example, strong external magnetic fields freeze in individual qubits, the final solution of the annealing process might not encode the ground state of the final Hamiltonian.
It is thus essential to prepare the system such that Many-Body Localizations are absent, and thus one must define a measure for MBL.

We note that this measure depends on the choice of partitions, which is, in principle, arbitrary.
To define a measure for MBL, the choice for partitions should encode the likelihood of frozen qubits.
Thus a natural choice for partitions are regions of different external magnetic fields.
For this choice, at the beginning of the anneal, the mutual information will be zero as all spins start orthogonal to the external field, and, in the end, the mutual information will be constant once a final solution is obtained.
Intermediate values for the mutual information encode the communication probability of different spin alignments.

Intuitively, to find a final state which globally minimizes the energy, we want the system to explore the entire state space.
That is, we want the system to have ergodicity in order to avoid Many-Body Localizations.
Thus the annealing process should be sufficiently slow to not introduce potential barriers which freeze in intermediate spin configurations.
On the other hand, decoherence in the hardware leads to the loss of quantum information.
It is, therefore, desirable to minimize the effect of relaxation and shorten the annealing process.
The interplay between these two competing effects can be observed by looking at the non-monotonous time dependency of the probability for reaching the global ground state.

@@

 and present an example implementation in case of the Minimal Dominating Set problem (\ref{sec:methods:mds}).
We review the annealing process on the utilized hardware (\ref{sec:methods:annealing}) and describe the computational framework which we used to simulate the annealing process (\ref{sec:methods:simulation}).
Section \ref{sec:results} summarizes our findings and presents...
We conclude this report in \ref{sec:conclusion}.
All algorithms used and data obtained are publicly available (\ref{sec:open-source})

@@

%----------------------------------------------------------------------------------------
\subsection{ILP QUBO}
\label{sec:results:qa}
%----------------------------------------------------------------------------------------

@@

For even the smallest graph $G(2)$, the MDS problem is not native to the chimera graph and must be embedded. Following the hypothesis of MBL, we therefore, must look at the values of $h_i$ after embedding. The qubits then split into two groups depending on the value of $h_i$ relative to $(\textrm{max}|\{h\}| + \textrm{min}|\{h\}|) / 2$ given the set of external magnetic fields $\{h\}$ defined by a specific embedding. Further detail is given in Sec.~\ref{sec:methods:minor_embedding} {\color{red} put in methods more details}. We study the effects offset the two groups of qubits over a total range between 0\% and 5\% of $c$. The results of this study is shown in Fig.~\ref{fig:dwave_offset}.

@@

%----------------------------------------------------------------------------------------
\subsection{Simulation Results}
%----------------------------------------------------------------------------------------

In order to understand the effects of annealing offsets, we simulate the annealing process for the $G(2)$ graph embedded in chimera topography.
In order to shorten the simulation time, we resample the DWave with a total annealing time of 1 $\mu s$.
The resulting ground state probability as a function of offset measured over 100,000 observations is shown in Fig.~\ref{fig:dwave1us}.

@@

%========================================================================================
\section{DISCUSSION}
\label{sec:discussion}
%========================================================================================

\subsection{Final remarks}
We would like to emphasize that while the annealing offsets are motivated by the MBL hypothesis, and the results also follow those expectations, we do not have definitive proof that MBL plays a crucial role.
The reason is because observations of MBL inevitably require the study of finite-size scaling, and our current simulation while being extremely thorough and explicit, is exponentially slow to evaluate making evaluations of even $G(3)$ untenable.
However, the intersection of time-dependent quantum mechanics and emergent phenomena such at MBL is an exciting direction that is pertinent to adiabatic quantum computing.

@@

%----------------------------------------------------------------------------------------
\subsection{Notes}
\label{sec:methods:ilp:notes}
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
\subsection{Quantum Annealing}
\label{sec:methods:annealing}
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
\subsubsection{Many-Body Localization}
\label{sec:methods:mbl}
%----------------------------------------------------------------------------------------

>>>>

\begin{abstract}
 While quantum computing proposes promising solutions to computational problems not accessible with classical approaches, due to current hardware constraints, most quantum algorithms are not yet capable of computing systems of practical relevance, and classical counterparts outperform them.
 To practically benefit from quantum architecture, one has to identify problems and algorithms with favorable scaling and improve on corresponding limitations depending on available hardware.
 For this reason, we developed an algorithm that solves integer linear programming problems (ILP), a classically NP-hard problem, on a quantum annealer (QA), and investigated problem and hardware-specific limitations.
 This work presents the formalism of how to map ILP problems to the annealing architectures.
 It illustrates the effects of decoherence and many-body localization, for the minimum dominating set problem, and compares annealing results against numerical simulations of the quantum architecture.
 We find that the QA ILP algorithm outperforms random guessing but is limited to small problems and that annealing schedules can be adjusted to reduce the effects of decoherence--which is qualitatively confirmed by simulations.
 {\color{red}Furthermore we find evidence for Many-body localization effects ... and find that FC decoherence is more relevant than local decoherence...}

@@

Integer linear programming (ILP) is an integer optimization problem subject to inequality constraints
\begin{align}
 \label{eq:initial-ip-def}
 \vec x_0 = \mathrm{arg} & \min\limits_{x}(\sum_i c_i x_i)
\end{align}
subject to
\begin{align}
 \label{eq:ilp-constraints}
  & \sum_i A_{ai}x_i +b_a \leq 0 \,, \\
  & x_i  \in \mathbbm{N}\, ,
\end{align}
where the natural numbers $\mathbbm{N}$ include zero, $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations. {\color{red} I don't actually understand this sentence.}

@@

An important ILP is the Minimum Dominating Set problem which is represantivly {\color{red} this is a typo, but i can't guess what word you're trying to use also...} considered in this work.
For a given a graph $G(E,V)$, defined by the set of $V$ vertices and $E$ edges, a dominating set $D$ is a specific subset of vertices $D \subseteq V$.
In particular $D$ is a dominating set if all vertices in $V$ but not in $D$ are adjacent to at least one vertex in $D$.
This is equivalent to requiring the set of nearest-neighbor vertices of $D$ (exclusive) and $D$ cover all vertices $N(D) \cup D = V$ (an example is given by Fig.~\ref{fig:dominating_sets}a).
The set $D$ is a minimal dominating set if there is no proper subset of $D$ that is a dominating set, {\it{i.e.}}, the removal of any vertex in $D$ results in $N(D) \cup D  \neq V$.
An example is given by Fig.~\ref{fig:dominating_sets}b.
The domination number of $D$ is given by the cardinality of $|D| \equiv \overline{\overline{D}}$.
The minimum dominating set (MDS) is defined by $D$ with the smallest domination number.
Fig.~\ref{fig:dominating_sets}c shows an example of the minimum dominating set of $G(V, E)$ and is different from the minimal dominating set.
We emphasize that while the maximum independent set is always a minimal dominating set as exemplified by Fig.~\ref{fig:dominating_sets}b, the minimum dominating set, in general, can have a smaller domination number.
As a result, the solution to the dominating set problem can not be obtained by solving for the maximum independent set, a problem that is well studied for quantum annealers~\cite{}.

@@

In this paper, we propose a new algorithm that maps ILP problems to QUBOs.
This is achieved by the introduction of slack variables $s$ which turn the inequalities eq.~\eqref{eq:ilp-constraints} to equalities
\begin{align}
  & \sum_i A_{a i}x_i + s_a + b_a = 0, \\
  & s_a \in \mathbb{N},                \\
  & x_i \in \mathbb{N}.
\end{align}
As a result, we can restrict both $x_i$ and $s_i$ to sample over only positive integer values. {\color{blue} In general, $s$ can be relaxed to span a fixed-point decimal number following Ref.~\cite{Chang:2018uoc} if the constraint equation is defined by non-integer coefficients.}

@@

If the annealer does not reach thermal equilibrium, for example, strong external magnetic fields freeze in individual qubits, the final solution of the annealing process might not encode the ground state of the final Hamiltonian. {\color{red} @CC is this statement actually correct?}
It is thus essential to prepare the system such that Many-Body Localizations are absent, and thus one must define a measure for MBL.

{\color{blue} These two paragraphs need to be changed (softened). The mutual information does not paint the story we need. Or more specifically, the MI we plot is not useful yet because we have no finite size scaling. In general, we do not have a good observable for MBL without finite size scaling. This is future work. The problem size is too small also to really see this effect. The best we can say is that we have this MBL hypothesis, and we set offsets with respect to that. So I think the paragraph above, and below should be changed to a short summary of Anderson Localization, and how our system is like a spin-glass, so we follow the AL logic. We then observe that the simulation agrees with DWave, and in the absense of unknown hardware effects like qubits freezing due to hardware limitations (this is a hardware problem actually, not a QM problem), that this scaling holds. So DWave claims offset is good to cure ``freezing'' problems cuz of hardware issues. We can claim this is more than hardware, but rooted in QM. This is the strongest statement we can make without getting skewered.}
{\color{red}
We note that this measure depends on the choice of partitions, which is, in principle, arbitrary.
To define a measure for MBL, the choice for partitions should encode the likelihood of frozen qubits.
Thus a natural choice for partitions are regions of different external magnetic fields.
For this choice, at the beginning of the anneal, the mutual information will be zero as all spins start orthogonal to the external field, and, in the end, the mutual information will be constant once a final solution is obtained.
Intermediate values for the mutual information encode the communication probability of different spin alignments.

Intuitively, to find a final state which globally minimizes the energy, we want the system to explore the entire state space.
That is, we want the system to have ergodicity in order to avoid Many-Body Localizations.
Thus the annealing process should be sufficiently slow to not introduce potential barriers which freeze in intermediate spin configurations.
On the other hand, decoherence in the hardware leads to the loss of quantum information.
It is, therefore, desirable to minimize the effect of relaxation and shorten the annealing process.
The interplay between these two competing effects can be observed by looking at the non-monotonous time dependency of the probability for reaching the global ground state.
}

@@

We first present the QUBO mapping for ILP (\ref{sec:results:qa1}), and demonstrate the methodology on an example implementation in case of the Minimal Dominating Set problem (\ref{sec:results:qa}) on the DWave quantum annealer. Results from quantum annealing are compared and discussed in contrast to simulations (\ref{sec:results:simulation}). All algorithms used and data obtained are publicly available (\ref{sec:open-source}).

@@

%----------------------------------------------------------------------------------------
\subsection{ILP QUBO}
\label{sec:results:qa1}
%----------------------------------------------------------------------------------------
{\color{red} This section and the methods section reads very disjoint. You also don't define everything here and then Reference the Methods for more information.}

@@

For even the smallest graph $G(2)$, the MDS problem is not native to the chimera graph and must be embedded. Following the hypothesis of MBL, we therefore, must look at the values of $h_i$ after embedding. The qubits then split into two groups depending on the value of $h_i$ relative to $(\textrm{max}|\{h\}| + \textrm{min}|\{h\}|) / 2$ given the set of external magnetic fields $\{h\}$ defined by a specific embedding. Further detail is given in Sec.~\ref{sec:methods:minor_embedding}. We study the effects offset the two groups of qubits over a total range between 0\% and 5\% of $c$. The results of this study is shown in Fig.~\ref{fig:baseline}.

@@

%----------------------------------------------------------------------------------------
\subsection{Simulation Results}
%----------------------------------------------------------------------------------------
\label{sec:results:simulation}
In order to understand the effects of annealing offsets, we simulate the annealing process for the $G(2)$ graph embedded in chimera topography.
In order to shorten the simulation time, we resample the DWave with a total annealing time of 1 $\mu s$.
The resulting ground state probability as a function of offset measured over 100,000 observations is shown in Fig.~\ref{fig:dwave1us}.

@@

\subsection{Final remarks}
\label{sec:results:final}
We would like to emphasize that while the annealing offsets are motivated by the MBL hypothesis, and the results also follow those expectations, we do not have definitive proof that MBL plays a crucial role.
The reason is because observations of MBL inevitably require the study of finite-size scaling, and our current simulation while being extremely thorough and explicit, is exponentially slow to evaluate making evaluations of even $G(3)$ untenable.
However, the intersection of time-dependent quantum mechanics and emergent phenomena such at MBL is an exciting direction that is pertinent to adiabatic quantum computing.

@@

%----------------------------------------------------------------------------------------
\subsubsection{Notes}
\label{sec:methods:ilp:notes}
%----------------------------------------------------------------------------------------

@@

@@

FC gives correct distribution. Local the simplest relaxation mode. Maybe run a plot showing we turn of local, and see same scaling. So offset is not directly tackling only this. Can show that in Fig. 7.

%----------------------------------------------------------------------------------------
\subsubsection{Many-Body Localization}
\label{sec:methods:mbl}
%----------------------------------------------------------------------------------------
{\color{red} I think we discussed enough about this in the main text. There's no need to talk about this here.}
